 

### 2021-08-01 vertical-thermal-review.Rmd
Leaf age section needs some more guidance
Working on Ecology section now

The vertical structure of microenvironments likely influences forest photosynthetic capacity and $T_{leaf}$ gradients [@rey-sanchez_spatial_2016] via leaf phenology. For example, deciduousness is an adaptation to conserve water during dry periods [@meakemRoleTreeSize2018], but is also structured across the vertical profile. Within tropical deciduous species, the fraction of seasonally deciduous individuals increases toward larger tree size classes in hotter, brighter environments, with understory individuals typically remaining evergreen [@conditQuantifyingDeciduousnessTropical2000c]. Even in evergreen forests, environmental variation alters leaf age distributions [@albertCrypticPhenologyPlants2019], which affects forest photosynthetic capacity [@wuLeafDevelopmentDemography2016; @niinemetsLeafAgeDependent2016] via age-specific leaf function [@albertCrypticPhenologyPlants2019]. Leaf age affects vulnerability to heat stress [@mariasImpactsLeafAge2017; @zhouLeafageEffectsTemperature2015] and can be expected to similarly scale through vertically structured phenology to affect forest responses to heat anomalies. For instance, in tropical evergreen forests, because leaf turnover rates of sun leaves in the upper canopy is faster than in the subcanopy, less damage accumulates [@millerOnlySunlitLeaves2021].

@@

Temperature can affect photosynthesis via direct and indirect pathways--directly, by altering photosynthetic enzyme activity and the electron transport chain, and indirectly through increased VPD causing stomatal closure [@lloydEffectsRisingTemperatures2008]. Photosynthesis has a peaked response to $T_{air}$, with the peak commonly found to corresponding to the prevalent ambient growing season conditions [@tanOptimumAirTemperature2017; @slotSituTemperatureResponse2017]. Beyond the optimum, photosynthesis decreases as a result of stomatal closure [e.g. @slotSituTemperatureResponse2017; @smithEmpiricalEvidenceResilience2020, temperate/boreal refs needed], and eventually due to biochemical constraints [@kumarathungeAcclimationAdaptationComponents2019a; @sageTemperatureResponseC32007; @varhammarPhotosyntheticTemperatureResponses2015].

>>>>

The vertical structure of microenvironments likely influences forest photosynthetic capacity and $T_{leaf}$ gradients [@rey-sanchez_spatial_2016] via leaf phenology. For example, deciduousness is an adaptation to conserve water during dry periods [@meakemRoleTreeSize2018], but is also structured across the vertical profile. Within tropical deciduous species, the fraction of seasonally deciduous individuals increases toward larger tree size classes in hotter, brighter environments, with understory individuals typically remaining evergreen [@conditQuantifyingDeciduousnessTropical2000c]. Even in evergreen forests, environmental variation alters leaf age distributions [@albertCrypticPhenologyPlants2019], which affects forest photosynthetic capacity [@wuLeafDevelopmentDemography2016; @niinemetsLeafAgeDependent2016] and heat stress via age-specific leaf function [@albertCrypticPhenologyPlants2019]. For instance, in tropical evergreen forests, because leaf turnover rates of sun leaves in the upper canopy is faster than in the subcanopy, less damage accumulates [@millerOnlySunlitLeaves2021]. Leaf age affects vulnerability to heat stress and has a strong effect on metabolism  [@mariasImpactsLeafAge2017; @zhouLeafageEffectsTemperature2015; @kikuzawaSynthesisRelationshipsLeaf2006], and can be expected to similarly scale through vertically structured phenology to affect forest responses to heat anomalies.

@@

Temperature can affect photosynthesis via direct and indirect pathways--directly, by altering photosynthetic enzyme activity and the electron transport chain, and indirectly through increased VPD causing stomatal closure [@lloydEffectsRisingTemperatures2008]. Photosynthesis has a peaked response to $T_{air}$, with the peak commonly found to corresponding to the prevalent ambient growing season conditions [@tanOptimumAirTemperature2017; @slotSituTemperatureResponse2017; @doughtyAreTropicalForests2008]. Beyond the optimum, photosynthesis decreases as a result of stomatal closure [e.g. @slotSituTemperatureResponse2017; @smithEmpiricalEvidenceResilience2020; @grossiordPlantResponsesRising2020; @fredeenTemperatureHumidityEffects1999], and eventually due to biochemical constraints [@kumarathungeAcclimationAdaptationComponents2019a; @sageTemperatureResponseC32007; @varhammarPhotosyntheticTemperatureResponses2015; @sharkeyEffectsModerateHeat2005].
